---
title: 'HW 20 -- Q2'
author: "Nick Huo"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        fig_width: 5.5
        fig_height: 3.5
        fig_caption: true
papersize: letter
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## a.

```{r}
fish <- read.csv("Fish.csv")
attach(fish)

mod_full <- lm(Weight ~ ., data=fish)
summary(mod_full)
```

It appears that `SpeciesParkki`, `SpeciesSmelt`, and `Length1` are useful as they have significant coefficients.

 - `SpeciesParkki`: The predicted weighted for Parkki fish is higher than that of Bream fish by 164.7 grams.
 
 - `SpeciesSmelt`: The predicted weighted for Smelt fish is higher than that of Bream fish by 446.1 grams.
 
 - `Length1`: The predicted weighted for a fish decrease by 80.3 grams for every 1 cm increase in vertical length.

## b.

```{r}
library(MASS)
mod_empty <- lm(Weight ~ 1, data=fish)
stepAIC(mod_empty, scope=list(lower=mod_empty, upper=mod_full),
        direction="both", k=log(length(fish)))
```

```{r, include=FALSE}
library(tidyverse)
library(caret)
library(leaps)

models <- regsubsets(Weight ~ ., data=fish, )
summary(models)
```

```{r, include=FALSE}
res.sum <- summary(models)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)

data.frame(
    Model = 1:8,
  Adj.R2 = res.sum$adjr2,
  CP = res.sum$cp,
  BIC = res.sum$bic
)
```

Using step-wise selection with BIC, the final model include `Species`, `Length1`, and `Length2` as predictors. This means that using a fish's species, vertical length, and diagonal length to predict its weight is a good approach.

## c.

```{r}
mod_final <- lm(Weight ~ Species + Length1 + Length2)
summary(mod_final)
```

I think the most important predictor in determining the weight of a fish is probably `Length2`. Because the diagonal length has the lower p-value and standard error for its coefficient, it has a stronger correlation with a fish's weight, compared to the correlation between the vertical length (`Length1`) and weight. And at the same time, the coefficient value of `Length2` has a higher magnitude, which might mean that it is more important, considering that the two lengths predictors values should have also be similar in magnitudes.


