---
title: "HW17 -- Q2"
author: "Nick Huo"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        fig_width: 5.5
        fig_height: 3.5
        fig_caption: true
papersize: letter
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q2

### Set-up

```{r}
ms <- read.csv("MuscleMass.csv")
ms$X_c <- ms$X - mean(ms$X)
attach(ms)
```

### a.

```{r}
fit <- lm(Y ~ X_c + I(X_c^2))
plot(X, Y, main="Age vs. Bone Mass (Quadratic Model Fitted)",
     xlab="Age", ylab="Bone Mass")
lines(sort(X), fitted(fit)[order(X)], col='red')
summary(fit)
```

From the scatter plot and the regression fit line, it appears that the model is fitting the data quite well. The model has $R^2=0.763$ and $R^2_{adj}=0.755$.


### b.

$H_0$: None of the predictors are significant.

$H_a$: At least one of the predictors are significant.

According to the regression model output, we see that the global F-test is significant as the p-value $<0.0001$ is smaller than $\alpha=0.05$ -- there is strong evidence suggesting at least one of the predictors is significant. We are able to reject $H_0$. Therefore, we conclude that at least one of the predictors is significant in predicting bone mass.


### c.

```{r}
newdata <- data.frame(X_c=48-mean(X))
predict(fit, newdata, level=0.95, interval="confidence")
```

The estimated mean muscle mass for women aged 48 years old is 99.25. 
We are 95% confident that the mean muscle mass for women aged 48 years old is between 96.28 and 102.22.

### d.

```{r}
predict(fit, newdata, level=0.95, interval="prediction")
```

The estimated muscle mass for a woman aged 48 years old is 99.25. 
We are 95% confident that muscle mass for a woman aged 48 years old is between 82.91 and 115.60.


### e.

$H_0$: The quadratic term is not significant.

$H_a$: The quadratic term is  significant.

The p-value of 0.081 suggests that there is some evidence suggesting the quadratic term is significant as it is larger than but close to $\alpha=0.05$. We failed to reject $H_0$. Thus, we conclude that the quadratic term is not significant and we can drop the quadratic term from the model. So the squared difference of age vs. the mean age is not significant in explaining bone mass.


### f.

```{r}
mean(X)
```

We know that $x=X-\bar{X}=X-59.98$.

Original Model: $\hat{Y}=82.96-1.18(x)+0.015(x)^2)$.
$$\hat{Y}=82.96-1.18(X-59.98)+0.015(X-59.98)^2)$$


Re-expressed Model: $\hat{Y}=82.96-1.18X+1.18*59.98+0.015(X^2-119.96X+3598))$
$$\hat{Y}=207.71-0.619X+0.015X^2$$


### g.

```{r}
cor(X, X^2)
plot(X, X^2)

cor(X_c, X_c^2)
plot(X_c, X_c^2)
```

The correlation coefficient between $X$ and $X^2$ is 0.996, which is very high.
The correlation coefficient between $x$ and $x^2$ is -0.038, very close to 0.
This suggests that it is useful to use a centered variable since it reduces the correlation between $X$ and $X^2$. It wouldn't be helpful to include both $X$ and $X^2$ as they are highly correlated. But if we use $x$ and $x^2$, because they are not correlated, they could explain additional variability in the response.


### h.

```{r}
library(MASS)
plot(fit, which=1)
plot(X_c, studres(fit), main="Studentized Residuals vs. X_centered",
     xlab="X_centered Values", ylab="Studentized Residuals")
plot(fit, which=2)
```

From the studentized residuals plots, we notice that the variance is not changing and there is no clear pattern. These suggest that constant variance and lineraity conditions are met. And looking at the QQ plot, we see that it mostly falls on the line, except at both tails, so it is approximately normal with a little bit of left-skewness. 


### i.

```{r}
fit <- lm(Y ~ X_c + I(X_c^2) + I(X_c^3))
plot(X, Y, main="Age vs. Bone Mass (Cubic Model Fitted)",
     xlab="Age", ylab="Bone Mass")
lines(sort(X), fitted(fit)[order(X)], col='red')
summary(fit)
```



$H_0$: None of the predictors are significant.

$H_a$: At least one of the predictors are significant.

According to the regression model output, we see that the global F-test is significant as the p-value $<0.0001$ is smaller than $\alpha=0.05$ -- there is strong evidence suggesting at least one of the predictors is significant. We are able to reject $H_0$. Therefore, we conclude that at least one of the predictors is significant in predicting bone mass.

Now if we look at the cubic term alone. We can test for its significance with

$H_0$: The cubic term is not significant.

$H_a$: The cubic term is significant.

The p-value of 0.72 suggests that there is little evidence suggesting the cubic term is significant as it is larger than $\alpha=0.05$. We failed to reject $H_0$. Thus, we conclude that the cubic term is not significant and we can drop the cubic term from the model. So the cubed difference of age vs. the mean age is not significant in explaining bone mass.

